An Embedding Framework for Consistent Polyhedral Surrogates

### 2022-05-10 poly-embed-journal/poly-embed-journal.tex
Small figure touchups and writing edits in 5.1

\tikzstyle{cell}=[dashed,thick]
\tikzstyle{simplex}=[thick]
\newcommand{\tikzfigscale}{2.2}
\newcommand{\placefiglabel}[1]{\node at (1.1,0,1.5) {#1};}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

\begin{definition}
  \label{def:calibrated}
  \jessiet{Need to be careful on use of $\gamma$.}
  \raft{Replace with prop[ell] or maybe define gamma}
  Let discrete loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
  If $(L, \psi)$ is calibrated with respect to $\ell$, we call $\psi$ a \emph{calibrated link.}
\end{definition}
It is well-known in finite-outcome settings that calibration is equivalent to \emph{consistency}, in the following sense (cf.~\citep{bartlett2006convexity,zhang2004statistical,agarwal2015consistent}).
Suppose we have the feature space $\X$ and label space $\Y$.
%For any data distribution $D \in \Delta(\X \times \Y)$, let $L^*$ be the best possible expected (over $D$) $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
We say a surrogate and link pair $(L,\psi)$ is consistent with respect to $\ell$ if, for all data distributions $D \in \Delta(\X \times \Y)$, and all sequences of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to the optimal surrogate loss $L^*$ (in expectation over $D$), the $\ell$-loss of the sequence $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to the optimal target loss $\ell^*$.
\bo{edited above slightly. I believe a technical definition of consistency would have a ``for all $\D$'', but we can probably skip that.}\jessie{Tried to make sure it had this technical correctness.}
\raf{Tried to clarify/correct the quantifiers.  OLD VERSION:
Let $L^*$ be the best possible expected $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
If $(L,\psi)$ is consistent, then for any sequence of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to $L^*$, the $\ell$-loss of $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to $\ell^*$ for all data distributions $D \in \Delta(\X \times \Y)$.}
%
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.
\jessiet{Commented out the following sentence as it was repetitive from earlier discussion. - Mar 24 22}

@@

Theorem \ref{thm:calibrated-separated} shows that calibration is equivalent to a geometric condition, which we call \emph{separation}, of a link function $\psi$.
Recall that for indirect elicitation, any point $u \in \Gamma(p)$ must link to a report $\psi(u) \in \gamma(p)$. \jessiet{We don't introduce indirect elicitation.  How do we want to address? - May 3, 22}
(In terms of losses, $u$ minimizing expected $L$-loss implies that $\psi(u)$ minimizes expected $\ell$-loss, with respect to $p$.)
The idea of separation is that points in the neighborhood of $u$ must also link to to a report in $\gamma(p)$.
Furthermore, there must be a uniform lower bound $\epsilon$ on the size of any such neighborhood.

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
We use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each such $u$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then there exists $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.
\jessiet{We write earlier that embedding is a special case of indirect elicitation, so a reader might be confused why (a) is not immediate.  I think this paragraph is saying that it is, but tries to provide intuition for why, right? - May 3, 22}

@@

\subsection{Applying the embedding framework}
When using our framework to study the (in)consistency of an existing surrogate $L:\reals^d \to \reals^\Y_+$, often the first step is determining the loss it embeds.
To this end, we suggest the following general approach.
First, for each $y\in\Y$, divide $\reals^d$ into a finite number of polyhedral regions on which $L(\cdot)_y$ is an affine function.
Second, identify the vertices of these polyhedral regions.%
\footnote{In some cases, these regions do not have vertices, such as the top-$k$ surrogates which are invariant in the all-ones direction; here one can restrict to a subspace, or otherwise select among equivalent reports.}
Third, conclude that the union of these vertices, $\Sc\subset\reals^d$, is a finite representative set for $L$.
% ~\citep[Statement 3.1.7]{grunbaum2013convex}.
Now $L$ embeds $L|_\Sc$ from Proposition~\ref{prop:representative-embeds-restriction}.
From here one can further remove redundant reports until one arrives at a tight embedding if desired.
Once the embedded discrete loss is known, the behavior of the surrogate becomes more clear: what problem it is solving, and for which restrictions on label distributions is it consistent for the original problem.

%% giving intuition for new figure
For example, \citet{ramaswamy2018consistent} propose the BEP surrogate, which embeds the discrete abstain loss (\S~\ref{sec:abstain}).
With the abstain loss in hand, it is easy to verify that the abstain-optimal report for any $p$ such that $\max_y p_y < 1/2$ is to ``abstain'' from predicting.
When juxtaposed with multiclass classification in Figure~\ref{fig:abstain-intuition-inconsistent}, one can observe that the BEP surrogate does not indirectly elicit, and therefore does not embed, the mode.
The intuition described by Figure~\ref{fig:abstain-intuition-inconsistent} additionally gives a procedure to validate inconsistency with respect to a target task.
\raf{Follow-up paragraph on prev sentence using abstain as an example}

@@

%% also interesting to use Sec 4 to verify proposed canonical links
\raft{Rework so emphasis is more on the general construction and less on testing a given link}
With an embedding in hand, Construction~\ref{const:eps-thick-link} provides a calibrated link function.
Yet for some target problems, the search for consistent surrogates has been restricted to those accommodating a particular canonical link function, such as $k$ largest coordinates of the surrogate report in top-$k$ classification (\S~\ref{sec:top-k}).
Interestingly, our construction is also useful in this situation, where one wishes to verify the consistence of a given proposed link $\psi$. 
Recall that Construction~\ref{const:eps-thick-link} produces a set-valued link envelope $\Psi$, which yields the possible values any $\epsilon$-separated link $\psi$ could map to.
If the given $\psi$ is indeed consistent, then it is $\epsilon$-separated for sufficiently small $\epsilon$, so one can always construct such a $\Psi$ and verify that $\psi(u) \in \Psi(u)$ for all $u\in\reals^d$.
More generally, while such canonical link functions may be intuitive for a given problem, our results suggest that researchers should consider setting them aside and instead let Construction~\ref{const:eps-thick-link} determine the link.
See \S~\ref{sec:lovasz-hinge} for a somewhat intricate example.

>>>>

\tikzstyle{cell}=[dashed,thick]
\tikzstyle{simplex}=[thick]
\newcommand{\tikzfigscale}{2.5}
\newcommand{\placefiglabel}[1]{\node at (1.1,0,1.5) {#1};}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

\begin{definition}
  \label{def:calibrated}
  Let discrete loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \prop{\ell}(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
  If $(L, \psi)$ is calibrated with respect to $\ell$, we call $\psi$ a \emph{calibrated link.}
\end{definition}
It is well-known in finite-outcome settings that calibration is equivalent to \emph{consistency}, in the following sense (cf.~\citep{bartlett2006convexity,zhang2004statistical,agarwal2015consistent}).
Suppose we have the feature space $\X$ and label space $\Y$.
%For any data distribution $D \in \Delta(\X \times \Y)$, let $L^*$ be the best possible expected (over $D$) $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
We say a surrogate and link pair $(L,\psi)$ is consistent with respect to $\ell$ if, for all data distributions $D \in \Delta(\X \times \Y)$, and all sequences of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to the optimal surrogate loss $L^*$ (in expectation over $D$), the $\ell$-loss of the sequence $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to the optimal target loss $\ell^*$.
\bo{edited above slightly. I believe a technical definition of consistency would have a ``for all $\D$'', but we can probably skip that.}\jessie{Tried to make sure it had this technical correctness.}
\raf{Tried to clarify/correct the quantifiers.  OLD VERSION:
Let $L^*$ be the best possible expected $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
If $(L,\psi)$ is consistent, then for any sequence of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to $L^*$, the $\ell$-loss of $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to $\ell^*$ for all data distributions $D \in \Delta(\X \times \Y)$.}
%
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.
\jessiet{Commented out the following sentence as it was repetitive from earlier discussion. - Mar 24 22}

@@

Theorem \ref{thm:calibrated-separated} shows that calibration is equivalent to a geometric condition, which we call \emph{separation}, of a link function $\psi$.
Recall that for indirect elicitation, any point $u \in \Gamma(p)$ must link to a report $\psi(u) \in \gamma(p)$. 
(In terms of losses, $u$ minimizing expected $L$-loss implies that $\psi(u)$ minimizes expected $\ell$-loss, with respect to $p$.)
The idea of separation is that points in the neighborhood of $u$ must also link to to a report in $\gamma(p)$.
Furthermore, there must be a uniform lower bound $\epsilon$ on the size of any such neighborhood.

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
We use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each such $u$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then there exists $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

@@

\subsection{Applying the embedding framework}
When using our framework to study the (in)consistency of an existing surrogate $L:\reals^d \to \reals^\Y_+$, often the first step is determining the loss it embeds.
To this end, we suggest the following general approach.
First, for each $y\in\Y$, divide $\reals^d$ into a finite number of polyhedral regions on which $L(\cdot)_y$ is an affine function.
Second, identify the vertices of these polyhedral regions.%
\footnote{In some cases, these regions do not have vertices, such as the top-$k$ surrogates which are invariant in the all-ones direction; here one can restrict to a subspace, or otherwise select among equivalent reports.}
Third, conclude that the union of these vertices, $\Sc\subset\reals^d$, is a finite representative set for $L$.
% ~\citep[Statement 3.1.7]{grunbaum2013convex}.
Now $L$ embeds $L|_\Sc$ from Proposition~\ref{prop:representative-embeds-restriction}.
From here one can further remove redundant reports until arriving at a tight embedding if desired.
Once the embedded discrete loss is known, the behavior of the surrogate becomes more clear: what problem it is solving, and for which restrictions on label distributions is it consistent for the original problem.

%% giving intuition for new figure
\raf{Follow-up paragraph on prev sentence using abstain as an example} \jessie{Attempted}
For example, \citet{ramaswamy2018consistent} propose the BEP surrogate, which embeds the discrete abstain loss (\S~\ref{sec:abstain}).
With the abstain loss in hand, it is easy to verify that the abstain-optimal report for any $p$ such that $\max_y p_y < 1/2$ is to ``abstain'' from predicting.
When juxtaposed with multiclass classification in Figure~\ref{fig:abstain-intuition-inconsistent}, one can observe that the BEP surrogate does not indirectly elicit the mode, as the level set $\prop{L^{BEP}}_{\vec 0} = \prop{\ell^{\mathrm{abs}}}_\bot \not \subseteq \mode_r$ for any $r$. \jessie{Notation}
As embedding is a special form of indirect elicitaion, we additionally know the BEP surrogate does not embed the mode.
The intuition described by Figure~\ref{fig:abstain-intuition-inconsistent} additionally gives a procedure to validate inconsistency with respect to a target task.

@@

%% also interesting to use Sec 4 for ``whole'' picture
\raft{Rework so emphasis is more on the general construction and less on testing a given link. \jessie{Attempted; old version in comments}}
With an embedding in hand, Construction~\ref{const:eps-thick-link} provides a calibrated link function.
% adding - May 10, 22
While the literature often ``guesses'' and verifies the calibration of propose links, Construction~\ref{const:eps-thick-link} removes the guesswork from the equation.
This is beneficial in when the most intuitive link functions are not calibrated and no known calibrated link is known; see \S~\ref{sec:lovasz-hinge} for a somewhat intricate example.
Moreover, as Construction~\ref{const:eps-thick-link} fully characterizes the set of possible calibrated link functions for a polyhedral embedding via the link envelope $\Psi$, one can also use $\Psi$ in Construction~\ref{const:eps-thick-link} to verify the calibration of proposed links, demonstrated in \S~\ref{sec:top-k}.
%% previous version - May 10 22
%Yet for some target problems, the search for consistent surrogates has been restricted to those accommodating a particular canonical link function, such as $k$ largest coordinates of the surrogate report in top-$k$ classification (\S~\ref{sec:top-k}).
%Interestingly, our construction is also useful in this situation, where one wishes to verify the consistence of a given proposed link $\psi$. 
%Recall that Construction~\ref{const:eps-thick-link} produces a set-valued link envelope $\Psi$, which yields the possible values any $\epsilon$-separated link $\psi$ could map to.
%If the given $\psi$ is indeed consistent, then it is $\epsilon$-separated for sufficiently small $\epsilon$, so one can always construct such a $\Psi$ and verify that $\psi(u) \in \Psi(u)$ for all $u\in\reals^d$.
%More generally, while such canonical link functions may be intuitive for a given problem, our results suggest that researchers should consider setting them aside and instead let Construction~\ref{const:eps-thick-link} determine the link.
%See \S~\ref{sec:lovasz-hinge} for a somewhat intricate example.

### 2022-05-10 poly-embed-journal/tikz/lovasz-links.tex
Small figure touchups and writing edits in 5.1

%\path[fill=softyellow, opacity = 0.1] (1.2,1.2) -- (1.2,-1.2) -- (-1.2,-1.2) -- (-1.2,1.2)  -- cycle;

\node at (1,1){\textbullet} (1,1)node[anchor= south west]{\scriptsize $ $};
\node at (1,0){\textbullet} (1,0)node[anchor= west]{\scriptsize $ $};
\node at (0,1){\textbullet} (0,1)node[anchor=south ]{\scriptsize $ $};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=north east]{\scriptsize $ $};
\node at (-1,0){\textbullet} (-1,0)node[anchor= east]{\scriptsize $ $};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south east]{\scriptsize $ $};

\node at (0,-1){\textbullet} (0,-1)node[anchor=north]{\scriptsize $ $};
\node at (1,-1){\textbullet} (1,-1)node[anchor=north west]{\scriptsize $ $};

\node at (0,0){\textbullet} (-.02,-.08)node[anchor=south west]{\scriptsize $ $};

@@

\node at (1,1){\textbullet} (1,1)node[anchor= south west]{\scriptsize $ $};
\node at (1,0){\textbullet} (1,0)node[anchor= west]{\scriptsize $ $};
\node at (0,1){\textbullet} (0,1)node[anchor=south ]{\scriptsize $ $};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=north east]{\scriptsize $ $};
\node at (-1,0){\textbullet} (-1,0)node[anchor= east]{\scriptsize $ $};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south east]{\scriptsize $ $};

\node at (0,-1){\textbullet} (0,-1)node[anchor=north]{\scriptsize $ $};
\node at (1,-1){\textbullet} (1,-1)node[anchor=north west]{\scriptsize $ $};

\node at (0,0){\textbullet} (-.02,-.08)node[anchor=south west]{\scriptsize $ $};

>>>>

\node at (1,1){\textbullet} (1,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,1)$};
\node at (1,0){\textbullet} (1,0)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,0)$};
\node at (0,1){\textbullet} (-0.03,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,1)$};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=south, minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,-1)$};
\node at (-1,0){\textbullet} (-1,0)node[anchor= south west, minimum height = 0.7cm, scale = 0.4]{\tiny $(-1,0)$};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,1)$};

\node at (0,-1){\textbullet} (-0.025,-1)node[anchor=south west,  minimum height = 0.75cm, scale = 0.3]{\tiny $(0, -1)$};
\node at (1,-1){\textbullet} (1,-1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1, -1)$};

\node at (0,0){\textbullet} (0,0)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,0)$};

@@

\node at (1,1){\textbullet} (1,1)node[anchor= south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,1)$};
\node at (1,0){\textbullet} (1,0)node[anchor= south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1,0)$};
\node at (0,1){\textbullet} (-0.03,1)node[anchor= south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,1)$};

\node at (-1,-1){\textbullet} (-1,-1)node[anchor=south, minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,-1)$};
\node at (-1,0){\textbullet} (-1,0)node[anchor= south, minimum height = 0.7cm, scale = 0.4]{\tiny $(-1,0)$};
\node at (-1,1){\textbullet} (-1,1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(-1,1)$};

\node at (0,-1){\textbullet} (-0.025,-1)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0, -1)$};
\node at (1,-1){\textbullet} (1,-1)node[anchor=south,  minimum height = 0.75cm, scale = 0.4]{\tiny $(1, -1)$};

\node at (0,0){\textbullet} (0,0)node[anchor=south west,  minimum height = 0.75cm, scale = 0.4]{\tiny $(0,0)$};
